---
title: "R_tutorial"
output: html_document
date: "2023-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(tidyverse)
```


# Basic datatypes
- numerical
- character
- logical
- vector/list
- matrix/dataframe

Numerical:
```{r}
x <- 5
y <- 3.14
print(x)
```

Numerical operations:
```{r}
x + y # 8.14
x - y # 1.86
x * y # 15.7
x / y # 1.5923567
```

Character:
```{r}
z <- "Hello, world !"

z <- paste("Hello,", "world", "!", sep = " ") # "Hello, world !"
```

Logical:
```{r}
a <- TRUE
b <- FALSE
```

Logical operations:
```{r}
a & b # TRUE AND FALSE -> FALSE 
a | b # TRUE OR FALSE -> TRUE
```

Vectors:
```{r}
my_vector<-c(1,2,6,7)

sequence<-seq(1,5,by=1) # c(1,2,3,4,5)
sequence<-1:5 # c(1,2,3,4,5)

c(sequence,my_vector) # c(1,2,3,4,5,1,2,6,7)
```

Lists:
```{r}
my_list <- list(a=5,b=8)
print(my_list[["a"]])
```

DataFrame:
```{r}
df<-data.frame(x=0:22,y=seq(3,3+22*2,by=2))
df$y<-sqrt(df$x)
df
```

Matrix:
```{r}
df.matrix<-as.matrix(df)
t(df.matrix)^2
```

# Functions/Loops
```{r}
power.function <- function(x,n){
  return(x^n)
}

power.function(2,4)
```

```{r}
for (i in seq(1,5)){
  print(power.function(2,i))
}
```

# Plotting
```{r}
ggplot(df,aes(x=x,y=y))+geom_line()
```

```{r}
ggplot(df,aes(x=x))+geom_boxplot()
```


# Statistics

# Data simulation
```{r}
set.seed(123)
n <- 1000
age <- rnorm(n, mean = 40, sd = 10)
gender <- sample(c("male", "female"), n, replace = TRUE, prob = c(0.4, 0.6))
smoking <- sample(c("never", "former", "current"), n, replace = TRUE, prob = c(0.5, 0.2, 0.3))
BMI <- rnorm(n, mean = 25, sd = 5)
confounder <- rnorm(n, mean = 20, sd = 10)
log_odds <- -0.5 + 0.05*age + 3.0*(gender == "male") + 0.3*(smoking == "current") - 0.1*BMI + 0.2*confounder
prob <- exp(log_odds) / (1 + exp(log_odds))
outcome <- factor(rbinom(n, 1, prob))
dataset <- data.frame(outcome, age, gender, smoking, BMI, confounder)
```


# Compare group means
```{r}
set.seed(123)
group1 <- rnorm(30, mean = 5, sd = 2)
group2 <- rnorm(30, mean = 7, sd = 2)

dat<-data.frame(group1=group1,group2=group2) %>% pivot_longer(cols=c("group1","group2"))

ggplot(data=dat,aes(x=value,color=name))+geom_boxplot()+theme_classic()
```
# T-test
```{r}
t.test(group1, group2)
```

# See whether outcome is related to gender
```{r}
table(dataset$gender, dataset$outcome)

ggplot(dataset,aes(x=gender,fill=outcome))+geom_bar(position="fill")
```

# Chisq Test
```{r}
chisq.test(dataset$gender, dataset$outcome)
```

# Compare age and smoking
```{r}
ggplot(dataset,aes(x=smoking,y=age))+geom_boxplot()+theme_classic()
```

# Run ANOVA to see whether age is related to smoking
```{r}
anova(model <- lm(age ~ smoking, data = dataset))
```

# Multivariable regression, controlling for other confounders
```{r}
summary(glm(outcome ~ age + gender + smoking + BMI + confounder, data = dataset, family = binomial))

```

# Machine learning
https://github.com/jlevy44/MolecularPathologyMachineLeaningTutorial/blob/main/R/1_fitting_ml_models.Rmd
